---
title: "311_Calls"
author: "David Rucinski"
date: "July 11, 2019"
output: 
    html_document:
    theme: cerulean
    highlight: espresso
    toc: true
    toc_float: false
    toc_depth: 4
    df_print: kable
    code_folding: hide
---

```{r include=FALSE}
knitr::opts_chunk$set(comment = NA)
```

## **Objective**
[Winnipeg](https://data.winnipeg.ca) has an open data catalogue, there are some options to visualize the data. It does not function well so I want to visualize it.

## **Summary**




## **Preparations**

*Packages used:*
```{r packages, message = FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(lubridate)
```

## **Analysis**

### Data

[Data](https://data.winnipeg.ca/Contact-Centre-311/311-Service-Request/4her-3th5) was extracted on 2019-06-30, by David Rucinski.

*Should I get API and update this?*
```{r Importing Data}

service_request <- fread("311_Service_Request.csv")
incoming_calls <-  fread("311_Incoming_Calls_By_Day.csv")


## API stuff
#
# https://github.com/Chicago/RSocrata Reference for API link in R
#
## Install the required package with:
## install.packages("RSocrata")
#
# library("RSocrata")
# 
# df <- read.socrata(
#   "https://data.winnipeg.ca/resource/4her-3th5.json",  #--> 311 service calls
#   app_token = "YOURAPPTOKENHERE",
#   email     = "user@example.com",
#   password  = "fakepassword"
# )
# https://data.winnipeg.ca/resource/lane_closure.json #--> Lane closure
#
#
# https://twitter.com/wpgpolice?ref_src=twsrc%5Egoogle%7Ctwcamp%5Eserp%7Ctwgr%5Eauthor #-->WPG Police twitter
# https://twitter.com/cityofwinnipeg #--> City of WPG twitter
# https://winnipeg.ctvnews.ca/ #--> ctv news 
# https://www.winnipegfreepress.com/ #---> free press
#  Also get weather

```


### Preprocessing

To do:
*Change chr as factors
*Clean date
*Time series with calls, any trends?
*Neighbourhood calls, who calls the most? (Way too many for comparing)
*Ward distribution for calls at time (x = time, y = n() , color = Ward) + geom_density 
```{r Factors}
glimpse(incoming_calls)
head(service_request)


service_request <- as.data.table(service_request)


service_request[, Neighbourhood:= as.factor(Neighbourhood)]
service_request[, Ward := as.factor(Ward)]
service_request[, `Service Area` := as.factor(`Service Area`)]
service_request[, `Service Request` := as.factor(`Service Request`)]

# Split Location 1, longitude and latitude -> change as numeric

service_request$Date <- parse_date_time(service_request$Date, '%m/%d/%Y %I:%M:%S %p')

head(service_request)
str(service_request)
levels(service_request$`Service Request`)

# "Frozen Catch Basin" sewer drains(catch basin)
```


```{r dates for incoming}

#Combine year, month, day --> put as date


```

#### Visualized 311 Service Requests
```{r, fig.width = 12, fig.height = 8}
service_request %>% filter(Date >= "2019-01-01 00:00:00 UTC") %>%
  group_by(`Service Request`) %>%
  summarize(sum = n()) %>%
  ggplot( aes(x = reorder(`Service Request`,sum), y = sum)) +
   geom_col( fill = "steelblue") +
   coord_flip() +
    geom_text(aes(label = round(sum), y = sum + 200))


#Where is garbage being missed most often?

service_request %>%
  filter(`Service Request` == "Missed Garbage Collection") %>%
  group_by(Neighbourhood) %>%
  summarize(sum = n()) %>%
  ggplot( aes(x = reorder(Neighbourhood, sum), y = sum)) +
   geom_col() +
   coord_flip()

# WAY TOO many neighbourhoods for that, though we could cut off to top 20?
# Maybe we should by ward

service_request %>%
  filter(`Service Request` == "Missed Garbage Collection") %>%
  group_by(Ward) %>%
  summarize(sum = n()) %>%
  ggplot( aes(x = reorder(Ward, sum), y = sum)) +
   geom_col() +
   coord_flip()


# Yes this one works better

```

**Remove seasonal calls, i.e. snow calls, for the current season**
```{r}
# To check this year filter(Date >= "2019-01-01 00:00:00 UTC") 

service_request %>% #filter(Date >= "2019-01-01 00:00:00 UTC") %>%
  group_by(`Service Request`) %>%
  filter(`Service Request` %in% c("Graffiti","Potholes", "Neighbourhood Liveability Complaint")) %>%
  count(Date) %>%
  group_by(`Service Request`, month=floor_date(Date, "month")) %>%
  summarize(amount=sum(n)) %>%
  ggplot( aes(x = month, y = amount, col = `Service Request`) ) +
  geom_line() #+
  #facet_wrap(~ `Service Request`)

#levels(service_request$`Service Request`)
```



